### 2022-05-16 main/chapter_2ljets.tex
add censored presel figures; begin event variables list

Diagrams illustrating the C1N2 and GMSB signal models we consider are shown in
Figure~\ref{fig:2ljets_signal_diagrams}.
Some example effects of these signals in signal regions are displayed in
Figure~\ref{fig:2ljets_signal_examples}.

At each parameter point, each signal model is added to the background model
and tested against data in the $\mathrm{CLs}$ prescription.
Contours showing which regions of signal model parameters are labelled as
excluded are shown for C1N2 in Figure~\ref{fig:2ljets_contours_c1n2} and for
GMSB in~\ref{fig:2ljets_contours_gmsb}.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figures/2ljets_summary_log.pdf}
\caption{%
Data of the $\twoljets$ electroweak analysis with \emph{post-fit} backgrounds
for which the lower panel shows $S_\mathrm{ATLAS}$ from
Equation~\ref{eqn:significance_atlas}.
Control, validation and signal regions are shown from left to right, with the
regions within each category ordered approximately by their typical $\ptmiss$.
Likelihoods from validation regions are not included in the fit.
The `Top' category contains $t\bar t$ and $tW$ processes, and
`Other' contains fake/non-prompt lepton, higgs, triboson, $t\bar tZ$, and other
rare top processes.%
}
\label{fig:2ljets_summary}
\end{figure}

@@

\section{Context}
% previous results on 2(/3)L from ATLAS
% other constraints on these models from ATLAS/CMS
% previous region design, motivation for this work

\section{Design}

@@

\section{Modelling}
% backgorund samples
% matrix method fakes

>>>>

We use two signal models in the $\twoljets$ electroweak analysis.
One is an MSSM chargino-neutralino model (C1N2) which acts to produce
$\chargino{1}\textrm{--}\neutralino{2}$ pairs which decay via $W$ and $Z$
channels respectively to jets and a lepton pair.
The other is a GMSB model which decays via Higgs and $Z$ bosons in variable
branching fractions.
Diagrams illustrating the C1N2 and GMSB signal models we consider are shown in
Figure~\ref{fig:2ljets_signal_diagrams}.
Some example effects of these signals in signal regions are displayed in
Figure~\ref{fig:2ljets_signal_examples}.

At each parameter point of a signal model, it is added to the background model
and tested against data in the $\mathrm{CLs}$ prescription.
Contours showing which regions where those tests label the signal model
parameters as excluded excluded are shown for
C1N2 in Figure~\ref{fig:2ljets_contours_c1n2} and for
GMSB in~\ref{fig:2ljets_contours_gmsb}.

@@

\begin{figure}[tp]
\centering
\includegraphics[width=\textwidth]{figures/2ljets_summary_log.pdf}
\caption{%
Data of the $\twoljets$ electroweak analysis with \emph{post-fit} backgrounds
for which the lower panel shows $S_\mathrm{ATLAS}$ from
Equation~\ref{eqn:significance_atlas}.
Control, validation and signal regions are shown from left to right, with the
regions within each category ordered approximately by their typical $\etmiss$.
Likelihoods from validation regions are not included in the fit.
The `Top' category contains $t\bar t$ and $tW$ processes, and
`Other' contains fake/non-prompt lepton, higgs, triboson, $t\bar tZ$, and other
rare top processes.%
}
\label{fig:2ljets_summary}
\end{figure}

@@

\clearpage
\FloatBarrier
\section{Context}
% previous results on 2(/3)L from ATLAS
% other constraints on these models from ATLAS/CMS
% previous region design, motivation for this work

\section{Event variables}
% lepton, jet definitions

Uncomfortably many event variables are used to define the regions of this
analysis.
Definitions of these variables are collected here.
All variables with a double object subscript (such as $\mll$, $\rjj$) refer to
the first two of those particles when ordered in decreasing $\pt$.
\begin{itemize}
\item $\etmiss$: missing transverse momentum.
\item $\etmisssig$: object-based missing transverse momentum significance.
It scales $\etmiss$ by an uncertainty derived from the resolutions of
reconstructed objects~\cite{atlas_met_significance}.
\item $\mll$: invariant mass of the hardest two leptons.
\item $\mjj$: invariant mass of the two hardest jets.
\item $\mbb$: invariant mass of the two hardest b-tagged jets.
\item $\mjetone$: mass of the hardest jet.
% TODO
% \item $\Rll$: estimated angular distance in $\eta,\phi$ between the lepton pair.
% \item $\Rjj$: estimated angular distance in $\eta,\phi$ between the two hardest jets.
% \item $\Delta\phi(p_{\ell\ell}, \Etmiss)$: estimated distance in azimuthal
% angle between the vector sum of the lepton pair and the missing transverse momentum.
% \item $\mttwo(\ell, \ell; 0)$: the `stransverse mass' between the two
% leptons and the missing transverse momentum.
% If two massive particles are pair produced and each decays to a
% lepton and something massless and invisible,
% then this estimates a greatest lower bound on the invariant
% mass of the pair produced particles.
% $\mttwo(\ell, \ell; 0)$ is useful for excluding standard model backgrounds;
% di-leptonic ttbar ($tt\rightarrow WW\rightarrow\ell\nu\ell\nu$)
% and $WW\rightarrow\ell\nu\ell\nu$ have a kinematic endpoint
% at the $W$ pole mass, and $\tau\tau$ is constrained within measurement error of zero.
% \item $\nbtag$: number of b-tagged jets with $\pt > 20$ GeV,
% where the tag is taken on the mv2c10 classifier at the $77\%$ working point.
% \item $n_{\mathrm{jets},30}$: number of reconstructed jets with $\pt > 30$ GeV.
% \item $\pt(j_1)$: transverse momentum of the hardest jet.
% \item $\Delta\phi(\mathrm{jet}_1,\etmiss)$: azimuthal angle between
% the hardest jet and missing transverse momentum vector estimates.
\end{itemize}
Most of these are physical properties of particles, or collection of particles.
Clearly, the stated values are not exact, but estimated values based on
approximate reconstructions.
While I think this distinction is important to remember, I have not spelled it
out on every line.
Brevity also has value.

\FloatBarrier
\section{Design}

The most impactful decision in a search is the design of its regions.
Like all decision in life, these designs must be made to balance conflicting
and vaguely-specified desires, and are informed with incomplete information.

This section describes the designs and rationales behind the control, validation
and signal regions of the $\twoljets$ analysis.

Splitting regions on $\etmisssig$ is the core idea behind the design.
The basic motivation for this design are displayed in
Figure~\ref{fig:2ljets_presel_met};
$\etmisssig$ does better than $\etmiss$ in separating background and signal
contributions.
But there is more --- $\etmisssig$ also does a decent job of separating signal
models from each other;
signals with larger mass differences in their decays to invisibles tend to have
larger $\etmisssig$.
Binning regions in $\etmisssig$ therefore also improves sensitivity to
individual parameter points and allows targeted selections on other, locally-relevant,
event variables.

\begin{figure}[tp]
\centering
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_logy.png}
\hfill
\includegraphics[width=0.49\textwidth]{figures/2ljets_presel_met_sig_logy.png}
\caption{
Illustration of how $\etmisssig$ beats $\etmiss$ alone in sensitivity to
example signal models.
At large $\etmisssig$, $t\bar t$ and other backgrounds are suppressed, leaving
quite pure diboson backgrounds and greater significance measures.
Signal models with smaller mass splittings tend to have the bulk of their
events at smaller $\etmiss$.
\\[0.5em]
This loose selection ``SRpreewk'' requires two same-flavor opposite-sign
signal leptons with
$\mll \in (71, 111)~\eV[G]$,
$\mjj \in (60, 110)~\eV[G]$,
$\nbtag \leq 1$,
$\etmiss > 100~\eV[G]$, and
$\etmisssig > 6$.
\\[0.5em]
``Significance'' displayed in the lower subplot uses the Binomial significance
$Z_{Bi}$ from~\cite{cousins2008evaluation}, with a $30\%$ background
uncertainty and yields taken to the right of a given value.
Large values indicate expected sensitivity to the signal model.
}
\label{fig:2ljets_presel_met}
\end{figure}

@@

\section{Modelling}
% TODO backgorund samples
% TODO matrix method fakes
% TODO GMSB reweighting
% TODO off-shell plus/minus split
% TODO generator filtering
